---
title: "Practical Machine Learning Course Project"
---
**Background (from the assignment)**  
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. 
 
In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 
 
**Overview**  
A training set of data is used to a predict the accelerometer data using a random forest classifier.  
 
**Loading libraries**  
```{r warning=FALSE, message=FALSE, }
library(AppliedPredictiveModeling)
library(caret)
library(rattle)
library(rpart.plot)
library(randomForest)
library(reshape2)
```
 
**Loading, cleaning, and reformatting data**  
```{r}
url_raw_training <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
file_dest_training <- "pml-training.csv"
url_raw_testing <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
file_dest_testing <- "pml-testing.csv"

train <- read.csv(file_dest_training, na.strings=c("NA",""), header=TRUE)
test <- read.csv(file_dest_testing, na.strings=c("NA",""), header=TRUE)

var <- names(train)[apply(train,2,function(x) table(is.na(x))[1]==19622)]   
train2 <- train[,var]
test2 <- test[,var[-length(var)]]

var2 <- melt(apply(train2,2,function(x) sum(ifelse(x=="",1,0)))==0)
select.var <- rownames(var2)[var2$value==TRUE]
train3 <- train2[,select.var]
test3 <- test2[,select.var[-length(select.var)]]
train4 <- train3[,names(train3[-c(1:7,length(train3))])]
test4 <- test3[,names(test3[-c(1:7)])]

correlations <- cor(train4)
highCorr <- findCorrelation(correlations, cutoff = .75)
predictor <- train4[, -highCorr]
filtered.test4 <- test4[, -highCorr]
classe <- train3$classe
trainData <- cbind(classe,predictor)
```
  
**Random forests**  
With 'classe' as the outcome variables, the remaining variables are used as predictor variables.    
```{r}
rfModel <- randomForest(classe ~ .,data = trainData,importance = TRUE,ntrees = 10)
```

**Confusion matrix and error rates**  
```{r}
confusion <- rfModel$confusion
sensitivity <- (confusion[2,2]/(confusion[2,2]+confusion[2,1]))*100
specificity <- (confusion[1,1]/(confusion[1,1]+confusion[1,2]))*100
overall_error <- rfModel$err.rate[length(rfModel$err.rate[,1]),1]*100
overall_accuracy <- 1-overall_error

out.test <- predict(rfModel,filtered.test4)

print(rfModel)
```

The unbiased estimate of the out of sample inaccuracy is 0.44%. 


**Submission code**
```{r}
answers <- as.vector(out.test)

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(answers)
```
